---
title: "Colorado Skip Year SGP Analyses"
subtitle: "A Historical Complarison of the 2019 SGP Results"
author:
  - name: Damian W. Betebenner
  - name: Adam R. VanIwaarden
  - name: <em>National Center for the Improvement</em>
  - name: <em>of Educational Assessment (NCIEA)</em>
date: September 2020
abstract_name: "Executive Summary"
abstract: "\\noindent DRAFT REPORT - DO NOT CITE! \\vspace{12pt} \\par The COVID-19 pandemic is having far reaching-effects on all facets of our lives. The impact on education included school closures in Spring 2020 and a 2020-2021 academic school year that likely will look quite different from pre-pandemic times. Colorado, like all states, is currently grappling with uncertainty regarding the summative spring assessments and accountability systems built from the results. Student academic growth is a significant component of Colorado's accountability system and the lack of Spring 2020 results has necessitated an investigation of what alternatives there are to the annual student academic growth calculations normally conducted. \\par This report summarizes findings of an investigation of using skip-year growth as a substitute for the typical annual (i.e., one-year) student growth calculation. To pursue this investigation, historical Colorado Measures of Academic Success (CMAS) data in ELA and Mathematics data was used in order to compare the SGPs derived from skip-year analyses (i.e., 2016 to 2018 and 2017 to 2019) versus one-year (non-skip year) analyses (i.e., 2017 to 2018 and 2018 to 2019). Comparisons between skip-year and one-year growth are examined at both the individual student level as well as the school level. \\par At the individual student level there is, not surprisingly, a very high correlation ($\\sim 0.85$) between skip-year SGPs and one-year SGPs (when the student has both). A very high correlation, however, does not preclude the existence of large differences for a substantial number of students. This is the case where more than 20 percent of the students have skip-year and one year SGP differences that exceed 15 in magnitude. \\par At the school level there is, again not surprisingly, a very high correlation ($\\sim 0.9$) between skip-year mean/median SGPs and one-year SGPs. However, the high correlation does mask approximately 10 percent of the schools whose mean/median SGP differs by 10 or more. The majority of schools demonstrating these large differences are elementary schools who have 4th grade SGPs when one-year SGPs are calculated but do not have 4th grade SGPs when skip-year SGPs are calculated because of the absence of a prior score. Thus, for those elementary schools, substantially different numbers of students are used to calculate the mean/median. As school grades are comprised of other quantities besides student growth, the overall impact on school grades is muted with only a handful of schools having their grades change using skip-year versus one-year student growth.\\par As spring 2021 approaches and state summative assessments are scheduled to be given in Colorado, the historical growth analyses discussed in this report support several recommendations: \\begin{itemize} \\item Historical analyses contained in this report do not unequivocally support or reject the use of skip-year student growth in lieu of one-year student growth for state accountability calculations at the school level. There is a high correlation between skip-year and one-year SGP results. However, there are a non-trivial number of schools where the differences are large enough to result in differing conclusions about student academic learning at the school.\\item Historical analyses contained in this report are done under much better educational circumstances than currently exist. We think that the current circumstances only adds to skepticism of using skip-year student growth in lieu of one-year student growth. Based upon current evidence, the best stance is to assume that skip-year growth is \\emph{not} a valid substitute for one-year growth in 2021 \\emph{unless} 2021 skip-year growth analyses align well with historical skip-year analyses. Given current disruptions to student education due to COVID-19 in Fall 2021, we do not expect the analyses will align well.\\item Even with analyses that align well in 2021, it is not clear whether business as usual accountability is politically tenable. The issue of summative testing in 2021 is only partly technical in nature. There is a large pragmatic, non-technical, component to consider. \\item Skip-year growth is extremely valuable to calculate even if not used for traditional accountability. Skip-year growth can be used to conduct investigations about the overall and differential impact of COVID-19 on the education of students. This has the potential to inform policy discussions about myriad of COVID-19 related issues including, for example, which modes of education best served students during the crisis.\\end{itemize}"
---

<!--SGPreport-->

<!-- 
This document was written by Damian Betebenner & Adam VanIwaarden for the State of Colorado Department of Education (CDE).

	Original Draft:  September 8, 2020
	Final Draft:     September 22, 202
-->


```{r, echo=FALSE, include=FALSE}
  ## set a universal Cache path
  knitr::opts_chunk$set(cache.path = "_cache/CO_Skip_Year")

  ##  Load some R packages and functions required for HTML table creation silently.  
  ##  Load SGP and other packages here to avoid messages.
  require(SGP)
	require(data.table)
	require(Gmisc)
  require(htmlTable)

  ##  Set Table, Figure and Equation Counters
  options(table_number=0)
  options("fig_caption_no"=0)
	options(fig_caption_no_sprintf = "**Figure %s:**   %s")
	options("fig_caption_no_roman"=FALSE)
	options("equation_counter" = 0)

	#   Colorado_Data <- copy(Colorado_SGP@Data)[VALID_CASE == "VALID_CASE" & YEAR == '2019' & CONTENT_AREA %in% c("ELA", "MATHEMATICS") & SCHOOL_ENROLLMENT_STATUS == "Enrolled School: Yes"]

  ###
	### Lists of CMAS subjects.  Used to subset summary tables and data below.
	###

  subject_order<-c("ELA", "MATHEMATICS")
	# EOCT_subjects<-c("ELA_PSAT_9", "ELA_PSAT_10", "ELA_SAT",
	# 								 "MATHEMATICS_PSAT_9", "MATHEMATICS_PSAT_10", "MATHEMATICS_SAT")
  
  ### Utility function - Assumes first two "levels" are the main dichotomy
  percent_demog <- function(demog) {
  	return(100*as.numeric(table(demog)[2])/as.numeric(sum(table(demog)[1:2])))
  }
  
  roundUp <- function(x, to=50) to*(x%/%to + as.logical(x%%to))

```

# Background

The COVID-19 pandemic is having far reaching-effects on all facets of our lives. The impact on education included school closures in Spring 2020
and a 2020-2021 academic school year that likely will look quite different from pre-pandemic times. Colorado, like all states, is currently grappling with
uncertainty regarding the summative spring assessments and accountability systems built from the results. Student academic growth is a significant component of Colorado's accountability system and the lack of Spring 2020 results has necessitated an investigation of what alternatives there are to the annual
student academic growth calculations normally conducted.

The most prominent alternative to the standard one-year student academic growth SGP utilized by Colorado is to calculate a two-year (skip-year) student
academic growth SGP from 2019 to 2021. The first question to such a proposal is, "Can you do that?" The answer is easy in one sense and not so easy
in another.

- In terms of performing the actual calculation, it is no more difficult to calculate skip-year growth than one-year growth. In fact, skip-year growth
is often calculated in states where students skip a grade (grade 8 to 10 growth) or where students take an end-of-course test and the most immediate
prior test is two years prior.
- Even though it is possible to calculate a skip-year growth quantity, it is not at all clear that it can be used in the manner that one-year growth
is used in state accountability systems.

Due to the many unknown aspects of school, testing, and accountability in the coming year, the best that any state can do at this point in time is
to examine historical data in anticipation of different situations that might arise in 2021. This report is part of the due diligence being conducted by
the Colorado Department of Education in preparation for Spring 2021.

```{r, cache=FALSE, results="asis", echo=FALSE}
	Literasee:::pageBreak()
```


#  Data

Colorado has extensive historical state summative assessment data from the [Colorado Measures of Academic Success (CMAS)](http://www.cde.state.co.us/assessment/cmas) 
in English language arts and mathematics for grades 3 to 8 to examine how skip-year growth and one-year growth are related. To do so we calculated 
skip-year growth for 2019 (using 2017 and 2016 as priors) and for 2018 (using 2016 and 2015 as priors). These skip-year growth quantities were then 
merged into files containing one-year SGPs permitting the comparison of skip-year and one-year growth at the individual level and, when aggregated 
by school, at the school level.

Table `r tblNum(1)` provides frequencies associated with student skip-year and one-year SGP counts by year, content area, and grade. The most
important aspect of Table `r tblNum(1)` is the fact that there are no 4<sup>th</sup> grade skip-year SGPs. This is due to the fact that there exists no
prior score with which to calculate growth with since there is no grade 2 exam. The absence of 4<sup>th</sup> grade SGPs has a significant impact on comparisons done
at the elementary school level. Since one-year growth summaries (mean or median) can include the 4<sup>th</sup> grade SGPs, one-year to skip-year comparisons
for elementary schools are not well aligned (using all growth data) since different grades exist. We discuss this issue later and supplement analyses
at the school level based upon only students having both a one-year and skip-year SGP.


```{r, cache=TRUE, echo=FALSE, include=FALSE}
	all.smry <- Colorado_Data[VALID_CASE=='VALID_CASE' & YEAR=='2019' & GRADE %in% 4:8, list(
									Mean_No_Skip = round(mean(NO_SKIP_SGP, na.rm = TRUE), 2),
									SD_No_Skip = round(sd(NO_SKIP_SGP, na.rm = TRUE), 2),
									Mean_Skip = round(mean(SGP, na.rm = TRUE), 2),
									SD_Skip = round(sd(SGP, na.rm = TRUE), 2),
									SGP_Corr = round(cor(NO_SKIP_SGP, SGP, use="pairwise.complete"), 3),
									Total_Students = .N,
									No_Skip_Count = sum(!is.na(NO_SKIP_SGP)),
									Skip_Count = sum(!is.na(SGP))), keyby = list(CONTENT_AREA, GRADE)]
	all.smry[, Difference := Skip_Count - No_Skip_Count]
	all.smry[, Pct_No_Skip := round((No_Skip_Count/Total_Students)*100, 1)]
	all.smry[, Pct_Skip := round((Skip_Count/Total_Students)*100, 1)]
	all.smry[, Difference_Pct := Pct_Skip - Pct_No_Skip]
	
	all.counts <- Colorado_Data[VALID_CASE=='VALID_CASE' & YEAR=='2019' & GRADE %in% 5:8, list( # only 5-8 -- possible to get both
									Total_Students = .N,
									No_Skip_Count = sum(!is.na(NO_SKIP_SGP)),
									Skip_Count = sum(!is.na(SGP)),
									Both = sum(!is.na(NO_SKIP_SGP) & !is.na(SGP)),
									Neither = sum(is.na(NO_SKIP_SGP) & is.na(SGP)),
									No_Skip_x_Skip = sum(!is.na(NO_SKIP_SGP) & is.na(SGP)),
									Skip_x_No_Skip = sum(!is.na(SGP) & is.na(NO_SKIP_SGP)))]
	both.pct <- round((all.counts$Both/all.counts$Total_Students)*100, 1)
	neither.pct <- round((all.counts$Neither/all.counts$Total_Students)*100, 1)
	no_skip.pct <- round((all.counts$No_Skip_x_Skip/all.counts$Total_Students)*100, 1)
	skip.pct <- round((all.counts$Skip_x_No_Skip/all.counts$Total_Students)*100, 1)
	# both.pct + neither.pct + no_skip.pct + skip.pct # shoul = ~100
```

```{r, results='asis', echo=FALSE, SGP_Counts}
	SGP_Counts <- all.smry[GRADE %in% 4:8, c(1:2, 8:10, 12:13)] # Percentages receiving SGPs
	SGP_Counts$CONTENT_AREA <- sapply(SGP_Counts$CONTENT_AREA, capwords, USE.NAMES=FALSE)
	SGP_Counts$CONTENT_AREA[duplicated(SGP_Counts$CONTENT_AREA)] <- ""
	SGP_Counts[, Total_Students := prettyNum(Total_Students, preserve.width = "individual", big.mark=',')]
	SGP_Counts[, No_Skip_Count := prettyNum(No_Skip_Count, preserve.width = "individual", big.mark=',')]
	SGP_Counts[, Skip_Count := prettyNum(Skip_Count, preserve.width = "individual", big.mark=',')]
	setcolorder(SGP_Counts, c(1:4, 6, 5, 7))
	
	row.names(SGP_Counts) <- NULL
	setnames(SGP_Counts, c("Content Area", "Grade", "Total Students", "Count", "Percent", "Count", "Percent"))
	# setnames(SGP_Counts, sapply(names(SGP_Counts), capwords, USE.NAMES=FALSE))
	SGP_Counts <- as.matrix(SGP_Counts)
	SGP_Counts[which(SGP_Counts==0)] <- '  --  '
	SGP_Counts[which(SGP_Counts==" 0.0")] <- ' -- '

  cat(dualTable(SGP_Counts, 
  	align=paste(rep('r', ncol(SGP_Counts)), collapse=''),
		n.cgroup=c(3, 2, 2), cgroup=c("", "No Skip", "Skip Year"),
		caption='No-skip and Skip-year SGP counts and percentages for 2019 by content area and grade.'))
```

```{r, results='asis', echo=FALSE, SGP_Counts_Alt}
	SGP_Counts <- all.smry[GRADE %in% 4:8, c(1:2, 8:10, 12:13)] # Percentages receiving SGPs
	SGP_Counts$CONTENT_AREA <- sapply(SGP_Counts$CONTENT_AREA, capwords, USE.NAMES=FALSE)
	SGP_Counts$CONTENT_AREA[duplicated(SGP_Counts$CONTENT_AREA)] <- ""
	SGP_Counts[, Total_Students := prettyNum(Total_Students, preserve.width = "individual", big.mark=',')]
	SGP_Counts[, `Count (Percent)` := 
						 	paste0(prettyNum(No_Skip_Count, preserve.width = "individual", big.mark=','), " (", format(Pct_No_Skip, drop0Trailing = FALSE), ")")]
	SGP_Counts[, ` Count (Percent)` := #  Need leading space in name to keep seperate...
						 	paste0(prettyNum(Skip_Count, preserve.width = "individual", big.mark=','), " (", format(Pct_Skip, drop0Trailing = FALSE), ")")]
	SGP_Counts[, c("No_Skip_Count", "Skip_Count", "Pct_No_Skip", "Pct_Skip" ) := NULL]
	row.names(SGP_Counts) <- NULL
	setnames(SGP_Counts, sapply(names(SGP_Counts), capwords, USE.NAMES=FALSE))
	SGP_Counts <- as.matrix(SGP_Counts)
	SGP_Counts[which(SGP_Counts=="0 ( 0.0)")] <- '   -- (--)   '

  cat(dualTable(SGP_Counts, 
  	align=paste(rep('r', ncol(SGP_Counts)), collapse=''),
		n.cgroup=c(3, 1, 1), cgroup=c("", "No Skip", "Skip Year"),
		caption='No-skip and Skip-year SGP counts and percentages for 2019 by content area and grade. VERSION 2'))
```


In general, there are fewer students with a skip-year SGP (prior from two years ago) than a one-year SGP. For all grades and content areas for which
both skip-year and one-year SGPs can exist, `r both.pct`% of students having both a one-year and skip-year SGP. `r no_skip.pct`% of students have a
one-year SGP but no skip-year SGP. Only `r skip.pct`% of students have a skip-year SGP but no one-year SGP. And `r neither.pct`% of students have neither.

Overall, based upon counts, there is not any systematic under-representation of students that would the representativeness of the group of students
used for skip-year growth calculations. For 2021, Table `r tblNum(1)` should have an additional column added indicating counts for students with
skip-year growth. If the values differ dramatically from those in 2019 then that would undermine claims about their representativeness. In
the next section we investigate the results more thoroughly to determine whether skip-year growth can be used in lieu of one-year growth.

```{r, cache=FALSE, results="asis", echo=FALSE}
	Literasee:::pageBreak()
```


#  Analyses and Results

Analyses were conducted for grades 4 to 8 for one-year SGPs and in grades 5 to 8 for skip-year SGPs in both ELA and mathematics for the 2019 CMAS administration. All analyses were conducted using the [`R` Software Environment](http://www.r-project.org/) [@Rsoftware] in conjunction with the
[`SGP` package](https://github.com/CenterForAssessment/SGP) [@sgp2020]. Source code associated with one-year analyses is available at
[this link](https://github.com/CenterForAssessment/Colorado) and source code associated with skip-year analyses is available at
[this link](https://github.com/CenterForAssessment/SGP_Research/tree/master/Colorado/Skip_Year_Analysis).

As mentioned previously, both sets of analyses are considered "standard" and resulted in no warnings or errors during calculation.
Goodness of fit plots associated with one-year and skip-year SGP analyses are provided in the appendix. In general, both one-year and skip-year
SGP analyses yielded excellent data fit. This is consistent with the excellent fit of analyses run previously in Colorado. Goodness of fit plots
for one-year and skip-year analyses 2019 for ELA and mathematics are available in Appendix A.
Because percentiles are uniformly distributed, mean, median and standard deviation for one-year and skip-year SGPs are practically identical.
In the following we discuss further results comparing one-year and skip-year SGPs at the individual level followed by results at the school level.

## Individual Level Results

One-year and skip-year SGPs are quantify growth for students across different time spans. For one-year SGPs, the time span is a single year. For
two-year SGPs the time span is two years. In general one would not communicate that a two-year SGP is the same as a one-year SGP because they
quantify something over a different time span.

As Colorado data analyses shows, however, the two quantities are highly similar. Correlations between one-year SGP and skip-year SGP by year,
content area, and grade ranged between `r round(min(all.smry$SGP_Corr, na.rm=TRUE), 2)`  and `r round(max(all.smry$SGP_Corr, na.rm=TRUE), 2)`. 
This indicates that though the quantities are for different time spans, those students showing high or low growth across one of the time spans 
often showed the same across the other time span. The results are consistent with those found in all other states examined thus far.

High correlations imply that the results demonstrate a strong linear relationship but do not imply that the results are identical or interchangeable.
Looking at individual differences between skip-year and one-year SGPs for students having both quantities shows that in 2019 in ELA and mathematics
for all grades approximately 20 percent of students showed differences between one-year and skip-year SGPs of at least 20 (or, equivalently, 80
percent had differences of less than 20). The results are not surprising in that an examination of student achievement scores over time show a
substantial number of students whose scores have wide fluctuations. Dropping the most recent score from growth calculations when it was in fact very
different will lead to highly discrepant SGPs.

Lastly, we examined whether there are systematic differences in terms of student achievement between the group of students with one-year SGPs but not
skip-year SGPs. Recall that students without a skip-year SGP are a small minority of students comprising approximately 5 percent of students in each 
grade and content area. In general, those students with no skip-year SGP (i.e., they do not have a prior from two-years earlier) were lower achieving
(z-score approximately -0.5 versus a z-score of 0.0 for all students). However, despite being lower achieving on average students with no-skip year
SGP didn't display any meaningful differences in terms of their academic growth (i.e., SGPs). It is unclear what based upon the data provided whether
there are systematic reasons for these students to not have a skip-year SGP (e.g., highly transient).

At the individual student level, the proposition of reporting skip-year (i.e., two-year) student growth in lieu of one-year student growth would
unnecessarily misrepresent what skip-year growth actually quantifies (growth over two years instead of a single year). The investigations carried
out with historical Colorado data suggest that there's a close relationship between the skip-year growth and one-year growth that could be used to
draw inferences about what the one-year growth would be if the student tested in 2020. In general, however, skip-year growth is the most relevant
and accessible indicator available and can serve many valuable investigative purposes apart from being a proxy for one-year growth.


## School Level Results

Of most interest to states in their investigation of skip-year growth is how well skip-year growth and one-year growth compare when aggregated to
the school level. Aggregation of SGPs to the school level is usually accomplished by taking the median or mean of the SGPs for the students
attending the school in the given year in each content area. One critical difference in skip-year growth and one-year growth is the absence of
SGPs for 4<sup>th</sup> grade students. A school with 4<sup>th</sup> graders would have student growth for those students included with one-year
growth but would not be included with skip-year growth. Elementary schools comprising grades K-5 or K-6 would have between a third and a half of
their students without skip-year growth when, previously, they had one year growth.

Because of disparities caused due to the fact that one-year analyses included 4<sup>th</sup> grade growth and skip-year analyses didn't, school 
level results were summarized in two ways:

1. Aggregates were calculated using all available skip-year and one-year growth scores.
2. Aggregates were calculated using only students with *both* a skip-year and one-year growth score.

The first method just takes the available growth scores (either skip-year or one-year) and calculates the mean/median from those. This method
would include 4<sup>th</sup> grade one-year SGPs for schools but those same schools would not have skip-year growth for those students. The second
method filters students to only those with both a one-year and skip-year SGP. This method omits 4<sup>th</sup> graders since they do not have a
skip-year SGP.


```{r, cache=TRUE, echo=FALSE, include=FALSE, School_Lev_Values}
##  Calculate School Demographic percentages if not already done
Colorado_Data[, PCT_SCHOOL_FRL := percent_demog(FREE_REDUCED_LUNCH_STATUS), by = c("YEAR", "CONTENT_AREA", "GRADE", "SCHOOL_NUMBER")]
Colorado_Data[, PCT_SCHOOL_ELL := percent_demog(ELL_STATUS), by = c("YEAR", "CONTENT_AREA", "GRADE", "SCHOOL_NUMBER")]
Colorado_Data[, PCT_SCHOOL_IEP := percent_demog(IEP_STATUS), by = c("YEAR", "CONTENT_AREA", "GRADE", "SCHOOL_NUMBER")]

##  Calculate Differences if not already done
Colorado_Data[, DIFF := as.numeric(NA)]
Colorado_Data[, DIFF := SGP - NO_SKIP_SGP]
Colorado_Data[, ABS_DIFF := as.numeric(NA)]
Colorado_Data[, ABS_DIFF := abs(SGP - NO_SKIP_SGP)]

###   Calculate School Level Values
sch.msgp.x <- Colorado_Data[VALID_CASE=='VALID_CASE' & YEAR == "2019" & !is.na(SCHOOL_NUMBER) & GRADE %in% 5:8,
              list(
                MEAN_NO_SKIP_SGP = mean(NO_SKIP_SGP, na.rm = TRUE),
                MEAN_SKIP_SGP = mean(SGP, na.rm = TRUE),
                COUNT_TOTAL = .N),
              keyby = c("CONTENT_AREA", "SCHOOL_NUMBER")][,DIFFERENCE_SKIP_NO_SKIP_MEAN_SGP:=round(abs(MEAN_SKIP_SGP-MEAN_NO_SKIP_SGP), 2)]

sch.msgp <- Colorado_Data[VALID_CASE=='VALID_CASE' & YEAR == "2019" & !is.na(SCHOOL_NUMBER),
              list(
                MEAN_NO_SKIP_SGP = mean(NO_SKIP_SGP, na.rm = TRUE),
                SD_NO_SKIP = sd(NO_SKIP_SGP, na.rm = TRUE),
                MEAN_SKIP_SGP = mean(SGP, na.rm = TRUE),
                SD_SKIP = sd(SGP, na.rm = TRUE),

                COUNT_NO_SKIP = sum(!is.na(NO_SKIP_SGP)),
                COUNT_SKIP = sum(!is.na(SGP)),
                COUNT_TOTAL = .N,

                MEAN_SCALE_SCORE = mean(SCALE_SCORE, na.rm=TRUE),
                MEAN_SCALE_SCORE_PRIOR = mean(SCALE_SCORE_PRIOR, na.rm=TRUE),
                MEAN_SCALE_SCORE_PRIOR_STANDARDIZED = mean(SCALE_SCORE_PRIOR_STANDARDIZED, na.rm=TRUE),
                MEAN_NO_SKIP_SGP_SCALE_SCORE_PRIOR_STANDARDIZED = mean(NO_SKIP_SGP_SCALE_SCORE_PRIOR_STANDARDIZED, na.rm=TRUE),
                PERCENT_FRED = PCT_SCHOOL_FRL[1],
                PERCENT_ELL = PCT_SCHOOL_ELL[1],
                PERCENT_IEP = PCT_SCHOOL_IEP[1]),
              keyby = c("CONTENT_AREA", "SCHOOL_NUMBER")][,DIFFERENCE_SKIP_NO_SKIP_MEAN_SGP:=round(abs(MEAN_SKIP_SGP-MEAN_NO_SKIP_SGP), 2)]

###   Calculate School Level Summaries

sch.diff.x <- sch.msgp.x[COUNT_TOTAL > 9, list(
		`Median Difference` = round(median(DIFFERENCE_SKIP_NO_SKIP_MEAN_SGP, na.rm=TRUE), 2),
		`95%ile Difference`=round(quantile(DIFFERENCE_SKIP_NO_SKIP_MEAN_SGP, probs=0.95, na.rm=TRUE), 2), N = .N), keyby = c("CONTENT_AREA")]

sch.cor.x <- sch.msgp.x[COUNT_TOTAL > 9, list(
    Skip_x_No_Skip = round(cor(MEAN_NO_SKIP_SGP, MEAN_SKIP_SGP, use="complete.obs"), 2), N = .N), keyby = c("CONTENT_AREA")]

sch.diff <- sch.msgp[COUNT_TOTAL > 9, list(
		`Median Difference` = round(median(DIFFERENCE_SKIP_NO_SKIP_MEAN_SGP, na.rm=TRUE), 2),
		`95%ile Difference`=round(quantile(DIFFERENCE_SKIP_NO_SKIP_MEAN_SGP, probs=0.95, na.rm=TRUE), 2), N = .N), keyby = c("CONTENT_AREA")]

sch.cor <- sch.msgp[COUNT_TOTAL > 9, list(
    Skip_x_No_Skip = round(cor(MEAN_NO_SKIP_SGP, MEAN_SKIP_SGP, use="complete.obs"), 2),
    Prior_Ach_x_No_Skip = round(cor(MEAN_NO_SKIP_SGP, MEAN_NO_SKIP_SGP_SCALE_SCORE_PRIOR_STANDARDIZED, use="complete.obs"), 2),
    Prior_Ach_x_Skip = round(cor(MEAN_SKIP_SGP, MEAN_SCALE_SCORE_PRIOR_STANDARDIZED, use="complete.obs"), 2),
    PCT_FRL_x_No_Skip = round(cor(MEAN_NO_SKIP_SGP, PERCENT_FRED, use="complete.obs"), 2),
    PCT_FRL_x_Skip = round(cor(MEAN_SKIP_SGP, PERCENT_FRED, use="complete.obs"), 2),
    PCT_ELL_x_No_Skip = round(cor(MEAN_NO_SKIP_SGP, PERCENT_ELL, use="complete.obs"), 2),
    PCT_ELL_x_Skip = round(cor(MEAN_SKIP_SGP, PERCENT_ELL, use="complete.obs"), 2),
    PCT_SWD_x_No_Skip = round(cor(MEAN_NO_SKIP_SGP, PERCENT_IEP, use="complete.obs"), 2),
    PCT_SWD_x_Skip = round(cor(MEAN_SKIP_SGP, PERCENT_IEP, use="complete.obs"), 2), N = .N), keyby = c("CONTENT_AREA")]

n.schools <- prettyNum(roundUp(length(unique(sch.msgp$SCHOOL_NUMBER))), big.mark=',')

```


Like with individual level results, correlations at the school level between one-year and skip-year median SGPs are very high. Table `r tblNum(1)`
reports the correlations (filtering on schools with 10 or more students) which approach 0.95 based upon approximately `r n.schools` Colorado schools.
The results demonstrate that those schools with high median one-year SGPs predominantly have high median skip-year SGPs and vice-versa. Though not
reported, similarly high correlations are observed at the district level as well as the classroom level.

```{r, results='asis', echo=FALSE, School_Lev_Cors}
	tmp.tbl <- data.table(`Content Area` = c("ELA", "Mathematics"), `All Students` = sch.cor$Skip_x_No_Skip, `Skip Subset` = sch.cor.x$Skip_x_No_Skip)

  cat(dualTable(as.matrix(tmp.tbl), 
  	align=paste(rep('r', ncol(tmp.tbl)), collapse=''),
		caption='School level correlations between median SGPs with unfiltered and filtered students.'))
```

Like at the individual level, a high correlation does not imply the quantities are identical or interchangeable. Absolute differences between
median SGPs were calculated for schools to examine the average magnitude of difference between one-year median SGP and skip-year median SGP. 
Table `r tblNum(1)` reports differences in school median SGP for all students and those students with both a skip-year and one-year SGP. 
On average the differences are small. However, as the 95th percentile shows, 5 percent of the school have report differences in median skip-year 
and one-year SGPs of nearly 7 in ELA and more than 9 in mathematics. From a practical standpoint, differences of more than 5 are associated with 
a small effect size. Moreover, a difference of 5 or more is likely to have some impact on the accountability designation that figures into the 
larger school rating. Overall, the differences are neither insignificant nor egregious.

```{r, results='asis', echo=FALSE, School_Lev_Diffs}
	tmp.tbl <- data.table(Filter = c("All Students", "", "Skip Subset", ""), rbindlist(list(sch.diff, sch.diff.x)))
	tmp.tbl[, N := prettyNum(N, big.mark=',')]
	tmp.tbl[, CONTENT_AREA := sapply(CONTENT_AREA, capwords)]
	setnames(tmp.tbl, "CONTENT_AREA", "Content Area")

  cat(dualTable(as.matrix(tmp.tbl), 
  	align=paste(rep('r', ncol(tmp.tbl)), collapse=''), n.rgroup=c(2, 2),
		caption='School level skip-year to one-year median and 95th percentile absolute differences.'))
```

School level differences like student level differences were, on average, minor. There were, however, numerous schools with skip-year/one-year
differences that were not minor and would possibly lead to a different accountability determination. We are currently working with the Colorado
Department of Education to re-run 2019 accountability calculations to determine the impact on school grades of using skip-year growth
in place of one-year growth. Those results will be added to this report as soon as they become available.

```{r, cache=FALSE, results="asis", echo=FALSE}
	Literasee:::pageBreak()
```

#  Summary

The analyses conducted in the report investigate whether skip-year SGPs can be used in place of the standard one-year SGPs. The report discussed
three potential issues:

- Are the students with skip-year SGPs systematically different than those with one-year SGPs?
- Are the one-year SGPs systematically different than the skip-year SGPs at the individual level?
- Are the one-year SGPs systematically different than the skip-year SGPs when aggregated to the school level?

\noindent For all the questions the answer is no. However, that is not to say that there weren't differences worth considering for a minority 
of students and schools. The results produced by these analyses do not disqualify the use of skip-year growth in lieu of one-year growth. 
However, they also do not qualify their use as well. Moreover, because the results here are generated from historical data derived under "usual"
educational circumstances, the current far from usual circumstances are likely to result in deviations that exceed what is reported here. 
Without confirmation in 2021 that skip-year deviations are no larger than they were historically, it seems imprudent at this point in time to 
confirm the use of skip-year growth for business as usual accountability.

Beyond the technical characteristics of skip-year growth lie numerous practical and political considerations on the use of skip-year growth for
accountability. As Colorado approaches 2021 any decisions related to student testing and accountability will require some technical considerations
but much more practical and political calculation to determine the best course going forward.

Assuming testing does occur in some form in 2021, it is highly recommended that growth be calculated regardless of whether it is used for
accountability.  There are numerous, valuable inquiries that can be addressed with skip-year SGPs available including:

- What was the overall impact on student learning in the state of Colorado due to COVID-19?
- What demographic subgroups fared best and which demographic subgroups fared most poorly in term of learning during the COVID-19 pandemic?
- Are there COVID-19 related instructional approaches that led to greater student learning than others? If so, what were they and how much more
learning were they associated with?

Critical to the investigation of COVID-19 related learning loss will be accurate data on COVID-19 related education programs that students
were assigned to. Having these data available will allow the Colorado Department of Education to fully take advantage of their student growth data
and enlighten stakeholders around the state as to what has occurred due to the COVID-19 pandemic.

```{r, cache=FALSE, results="asis", echo=FALSE}
	Literasee:::pageBreak()
```

# References
